---
title: "p8105_hw2_kw3180"
author: "Kino Watanabe"
date: "2025-09-26"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```
# Question 1

Clean the data in pols-month.csv:

```{r}
pols_month_df = 
    read_csv("fivethirtyeight_datasets/pols-month.csv") |> 
    janitor::clean_names() |> 
    separate(mon, into = c("year", "month", "day"), convert = TRUE) |> 
    mutate(month = month.name[month],
      president = prez_gop - prez_dem,
      president = case_when(
          president > 0 ~ "gop",
          president < 0 ~ "dem"),
      year = as.numeric(year)) |> 
    select(-prez_gop, -prez_dem, -day)
```

Clean the snp.csv data
```{r}
snp_df = 
    read_csv("fivethirtyeight_datasets/snp.csv") |> 
    janitor::clean_names() |> 
    separate(date, into = c("month", "day", "year"), convert = TRUE) |> 
    mutate(month = month.name[month],
           year = case_when(
             year <= 15 ~ year + 2000,
             year > 15 ~ year + 1900
           )) |> 
    select(year, month, close, -day)
```

Clean the unemployment data

```{r}
unemployed_df = 
    read_csv("fivethirtyeight_datasets/unemployment.csv") |> 
    janitor::clean_names() |> 
    pivot_longer(
      cols = jan:dec,
      names_to  = "month",
      values_to = "unemployment_percent"
    ) |> 
    mutate(
        month = month.name[match(str_to_title(month), month.abb)]
    )
```

Merge pols_month_df and snp_df

```{r}
pols_snp_df = 
    left_join(pols_month_df, snp_df, by = c("year", "month")) |> 
    select (year, month, president, close, everything())
    
```
Merge unemployed_df with pols_month_df

```{r}
pols_snp_unemploy_df = 
    left_join(pols_snp_df, unemployed_df, by = c("year", "month")) |> 
    select (year, month, president, close, unemployment_percent, everything())
```


# Question 2 

Clean Mr. Trash Wheel
```{r}
mrtrash_df = 
    read_excel("202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Mr. Trash Wheel",
                      range = "A2:N653",
                      na = c("na", ".", "")) |> 
    janitor::clean_names() |> 
    mutate(
      sports_balls = as.integer(round(sports_balls)),
      data = "mrtrash",
      year = as.numeric(year)) |> 
      select(month, year, date, data, everything(), -dumpster)
```

Clean Professor Trash Wheel
```{r}
prof_trash_df =
    read_excel("202409 Trash Wheel Collection Data.xlsx",
                    sheet = "Professor Trash Wheel",
                    range = "A2:M120",
                    na = c(".", "NA", "")) |> 
    janitor::clean_names() |> 
    mutate(
        data = "prof_trash"
    ) |> 
    select(month, year, date, data, everything(), -dumpster)
```

Clean Gwynnda Trash Wheel
```{r}
gwynnda_df = 
    read_excel("202409 Trash Wheel Collection Data.xlsx",
                  sheet = "Gwynnda Trash Wheel",
                  range = "A2:L265",
                  na = c(".", "NA", "")) |> 
    janitor::clean_names() |> 
    mutate(
        data = "gwynnda"
    ) |> 
    select(month, year, date, data, everything(), -dumpster)
```

Bind the data sets together

```{r}
bind_trash_df = 
    bind_rows(mrtrash_df, prof_trash_df, gwynnda_df) |> 
    janitor::clean_names() |> 
    arrange(date)
```

* In the combined data set of Mr. Trash, Professor Trash, and Gwynnda called bind_trash_df, there are `r nrow(bind_trash_df)` observations and `r ncol(bind_trash_df)` variables. The key variables include the times of data collection (`month`, `year`, `date`), the trash wheel data set identifier (`data`), the weight of trash in tons (`weight_tons`), and trash counts (`plastic_bottles`, `polystyrene`, `cigarette_butts`, `glass_bottles`, `plastic_bags`, `wrappers`, and `sports_balls`).

* The total weight of trash collected by Professor Trash Wheel is `r bind_trash_df |> filter(data == "prof_trash") |> summarise(total_weight = sum(weight_tons, na.rm = TRUE))|> pull(total_weight)` tons.

* The total number of cigarette butts collected by Gwynnda in June of 2022 is
`r bind_trash_df |>  filter(data == "gwynnda", year == 2022, month == "June") |>  summarise(cignum = sum(cigarette_butts, na.rm = TRUE)) |> pull(cignum)` butts.


# Question 3

Load Zip Codes
```{r}
zip_codes_df = 
    read_csv("zillow_data/Zip Codes.csv", na = c(".","NA","")) |> 
    janitor::clean_names() |> 
    mutate(county = 
             case_match(county,
                        "New York" ~ "Manhattan", .default = county)) |> 
    select(county, zip_code, neighborhood, everything())
    
```

Load Zillow Observed Rent Index (ZORI) in New York City between January 2015 and August 2024
```{r}
zori_df = 
      read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv", 
               na = c(".", "NA", "")) |> 
      janitor::clean_names() |> 
      rename(
          zip_code = region_name,
          county = county_name) |> 
      pivot_longer(
         cols = x2015_01_31:x2024_08_31,
         names_to = "date",
         values_to = "rental_price",
        names_prefix = "x") |> 
      mutate(county = replace(county, county == "New York County", "Manhattan"),
            county = str_replace(county, " County", ""),
            zip_code = as.numeric(zip_code),
            date = as.Date(str_replace_all(date, "_", "-"), format = "%Y-%m-%d")) |> 
        select(county, zip_code, date, rental_price, everything(),-state_name) |> 
        arrange(date)
```

We want to priortize zori_df with the date and price information. We will now merge zori_df and zip_codes_df.

```{r}
zip_zori_df = 
  left_join(zori_df, zip_codes_df, by = c("zip_code", "county")) |> 
  select(zip_code, county, neighborhood, date, rental_price, everything())
```




